### 2024-02-15 outputs/paper/paper.qmd
polished paper

---
title: "The Power of Liberty in the American Workplace: Unveiling the Global Impact of Free Will Perceptions on Job Satisfaction"
author: Sami El Sabri and Liban Timir
thanks: "Code and data are available at: https://github.com/samielsabri/freewill_analysis. SSRP replication available at: https://www.socialsciencereproduction.org/reproductions/94caaeaf-7e11-4efa-9269-7729fc2e0837/index"
date: 02-13-2024
date-format: long
abstract: "This study examines the relationship between belief in free will and job satisfaction in the United States, using data from Amazon Mechanical Turk and the World Values Survey. It replicates and extends previous findings by Feldman et al. (2017), confirming that free will belief positively correlates with job satisfaction in both samples. The analysis identifies job autonomy as a key mediator in this relationship. Results highlight the significant role of free will beliefs in enhancing workplace satisfaction and emphasize the importance of considering cultural and societal factors particular to the American context"
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(ggplot2)
library(MASS)
library(tidyverse)
library(car)
library(psych)
library(knitr)
library(kableExtra)
library(corrplot)
library(countrycode)
library(purrr)

wvs_data_usa <- read_csv(here::here('outputs/data/wvs_data_usa.csv'))
wvs_filtered <- read_csv(here::here("inputs/data/study_3/wvs_filtered.csv"))
satisfaction_data_usa <- read_csv(here::here('inputs/data/study_2/FW satisfaction-Study 2-data.csv'))
satisfaction_data_usa_rel <- read_csv(here::here('outputs/data/satisfaction_data_usa_rel.csv'))

```

# Introduction

In recent years, the intersection between personal beliefs in free will and cognitive behavioral outcomes has emerged as a pivotal area of psychological inquiry. While the existence of free will is a highly theoretical and philosophical question, the belief in free will has undeniable effects on various other psychological constructs such as autonomy, meaningfulness, and self-efficacy (@baumeister2014, @alquist2013). Despite the robustness of these relationships, there remain gaps in the literature of the specific dynamics of these effects and the generalizability in field settings. @feldman2017 conducted a study investigating the effects of the belief in free will on job satisfaction and attempted to draw cross-cultural and longitudinal conclusions on the positive correlation between the belief in free will and job satisfaction.

They executed three distinct studies: the first with real estate agents in Taiwan over three months, the second involving American workers over six months using Amazon Mechanical Turk, and the third utilizing a large cross-cultural sample from the World Values Survey encompassing 16 countries. This multifaceted approach allowed them to assess the generalizability of free will beliefs as predictors of job satisfaction across various cultural contexts and professional environments, providing a robust examination of how these beliefs influence job satisfaction over time.

Utilizing the same dataset and a similar methodology to @feldman2017, we aim to replicate their results in the context of the United States, a very heterogenous country that is characterized by a high importance of individualist, but also cultural diversity and socioeconomic inequality. While Feldman et al. did investigate the relationship between the belief in free will and job satisfaction in the United States, they did not integrate these results to the cross-cultural analysis, as the US were not included in the sample of 16 countries. This paper, therefore, seeks to fill this gap and reproduce these aspects of the Feldman et al. paper: (1) the relationship between the belief in free will and job satisfaction applies to the context of American adults working at Amazon Mechanical Turk (2) free will is a significant predictor of job satisfaction at the national-level, using the World Values Survey. We will also consider whether there are any significant differences between the national-level data and the Amazon Mechanical Turk data. The estimand is the correlation between the belief in free will and job satisfaction. Utilizing Amazon Mechanical Turk and World Values Survey data, we found that belief in free will significantly correlates with job satisfaction both at individual and national levels. Additionally, we found job autonomy as a critical mediator in this relationship, while other agency constructs and demographic factors play a less important role in impacting workplace satisfaction across diverse American contexts. We also found American participants to be among the world's highest believers in free will, further emphasizing the importance of considering the local context.

We begin our paper with a closer look at the data source and measurement (@sec-data), followed by a selection of relevant results (@sec-results). Finally, we will conclude with a discussion of our findings to integrate our replication with the original paper by @feldman2017 and the broader literature on free will beliefs (@sec-discussion). We will also discuss potential ethical biases, limitations, and future directions. Our reproduction was conducted using the statistical programming language R [@citeR]. In the data analysis and visualization process, we also made use of the following R packages: tidyverse [@citetidyverse], dplyr [@citedplyr], knitr [@citeknitr], kableExtra [@citekExtra], purrr [@citepurrr], psych [@citepsych], car [@citecar], MASS [@citeMASS], and countrycode [@citecc].

# Data {#sec-data}

## Source {#sec-source}

For the purposes of this paper, we make use of two distinct data sources.

The first data source is original survey data conducted by @feldman2017 on Amazon Mechanical Turk (MTurk), a popular online platform connecting "requesters" who offer tasks with "workers" who complete them for payment. The platform hosts a diverse, global workforce engaging in a range of tasks, from simple questionnaires to complex assignments requiring specialized skills. Compensation varies widely, with the average task priced at several dollars for about half an hour's work. This study surveyed American MTurk workers across two waves about their job satisfaction on MTurk, reflecting their overall work experience on the platform. The initial survey included 209 participants, with a follow-up six months later achieving a 66% response rate. @feldman2017 only included American workers in their study, making this data source appropriate for this investigation. The raw dataset comprised 143 variables, including both demographic variables and survey items relating to the outcomes and controls of interest.

The second data source is the World Values Survey (@citewvs), which is a large cross-cultural and cross-occupational survey collected in waves between 1981 and 2022 of 450,869 participants from over 40 countries. @feldman2017 used the WVS of 2008 and included a sample of 16 countries. In our own analysis, we used the most recent dataset and filtered for the same sample of 16 countries, as well as for data on the United States of America. These filtered datasets comprised 136,171 and 12,983 participants, respectively and only included a small subset of nine the survey questions (including demographic information).

## Measurement {#sec-measurement}

### Belief in Free Will

@feldman2017 measured belief in free will based on an adapted 9-item personal agency and free will subscale of the "Free Will and Determinism Scale" [@rakos2008]. Participants indicated their responses from 1 (Strongly disagree) to 6 (Strongly Agree) on the following nine items:

1.  I have free will
2.  Free will is a part of the human spirit
3.  Free will is a basic part of human nature
4.  I have free will even when my choices are limited by external circumstances
5.  People have free will regardless of wealth or life circumstances
6.  Life's experiences cannot eliminate a person's free will
7.  I am in charge of the decisions I make
8.  I decide what action to take in a particular situation
9.  I am in charge of my actions even when my life's circumstances are difficult

As seen in @tbl-reliability, the reliability coefficient for this subscale indeed measuring the belief in free will is $\alpha$ = 0.88. The full scale by @rakos2008 additionally includes psychological constructs related to moral and personal responsibility, higher power control etc., but these were deemed as weakening the reliability of this measure [@feldman2017].

In the WVS, we followed the assumption [@clark2014, @feldman2017] that the following item (A173 in the raw survey dataset) accurately assesses belief in free will: "“Indicate how much freedom of choice and control you feel you have over the way your life turns out” *(1 = no choice and control to 10 = a great deal of choice and control)*. This WVS item was previously used as a measure of the belief in free will [@clark2014]

### Job Satisfaction
In @feldman2017, job satisfaction was measured using a 5-item short scale version of @brayfield1951, constructed by @judge1998. Participants indicated their responses from 1 (Strongly disagree) to 7 (Strongly Agree) on the following five items:

1. I feel fairly well satisfied with my present job doing MTurk tasks
2. Most days I am enthusiastic about the work I do on MTurk
3. Every time I work on tasks on MTurk it feels like forever (reversed)
4. I find real enjoyment in the work I do on MTurk
5. I consider the kind of work I do on MTurk rather unpleasant (reversed)

Items 3 and 5 were inverted as a data cleaning step in order to correctly calculate correlations. Job Satisfaction was measured at two time points (three months apart) by @feldman2017, but at both times the reliability coefficients were fairly high ($\alpha$ = 0.79 and $\alpha$ = 0.84), indicating that these five items indeed measure job satisfaction among MTurk workers in the sample.

In the WVS, we followed @feldman2017 and @dobrow2016 and used the following item (C033): “Overall, how satisfied or dissatisfied are you with your job?” *(1 = dissatisfied to 10 = satisfied)*.

### Job Autonomy
Job Autonomy was measured using a 3-item scale by @hackman1980, which @feldman2017 adapted for MTurk. Participants indicated their responses from (Strongly disagree) to 7 (Strongly Agree) on the following three items:

1. I have significant autonomy in determining how I do my job on MTurk
2. I can decide on my own how to go about doing my work on MTurk
3. I have considerable opportunity for independence and freedom in how I do my job on MTurk

@feldman2017 measured job autonomy also at two time points, with comparably high reliability coefficients ($\alpha$ = 0.83 and $\alpha$ = 0.85), also supporting the assumption that the above scale measures job autonomy accurately.

In the WVS, we directly replicated @feldman2017 and used the following single-item survey question (C034): “How free are you to make decisions in your job?” *(1 = none at all to 10 = a great deal)*.

### Control Variables
In the MTurk dataset, we included all of @feldman2017 control variables (locus of control, self-control, self-esteem, self-efficacy, and implicit beliefs), which are all measured with similarly phrased survey questions as the aforementioned variables of interest. See the appendix (@sec-apx-controls) for detailed survey questions. As shown in @tbl-reliability, we were able to successfully replicate the coefficients for most constructs. The only deviation is with the trait locus of control, which has a coefficient $\alpha$ = 0.58 in @rotter1966, but 0.61 in our replication, indicating even greater confidence that the survey questions indeed measure the underlying construct.

For their analysis of the WVS data, @feldman2017 did not include any control variables. Since we were interested in understanding more about the relationships between the variables at the national level, we included the following two variables:

* Belief in Fate: We expect this variable to negatively correlate with belief in free will, such that it can serve as some indication of internal validity and reliability. Not all waves of WVS data asked this question, but where possible we used the following question (F198):
  + Some people believe that individuals can decide their own destiny, while others think that it is impossible to escape a predetermined fate. Please tell me which comes closest to your view on this scale on which 1 means “everything in life is determined by fate,” and 10 means that “people shape their fate themselves.”
  + We inverted this scale, such that 1 means "people shape their fate themselves.” and 10 means “everything in life is determined by fate”, making it more consistent with the ordering logic of other variables used.

* Pride in Job: We expect this variable to positively correlate with job satisfaction and job autonomy, such that it can serve as some indication of internal validity and reliability. Not all waves of WVS data asked this question, but where possible we used the following question (C031):
  + How much pride, if any, do you take in the work that you do? (1 = "A great deal", 3 = "Little").
  + We also inverted this scale, such that 1 means "Little" and 3 means "A great deal", making it making it more consistent with the ordering logic of other variables used.

```{r, include=FALSE, message=FALSE, echo=FALSE}
js1_items <- satisfaction_data_usa[, c("jobsat_1", "jobsat_2", "jobsat_3", "jobsat_4", "jobsat_5")]
alpha_js1 <- alpha(js1_items, check.keys=TRUE)
alpha_js1 <- alpha_js1$total$raw_alpha
alpha_js1 <- round(alpha_js1,2)

js2_items <- satisfaction_data_usa[, c("jobsat2_1", "jobsat2_2", "jobsat2_3", "jobsat2_4", "jobsat2_5")]
alpha_js2 <- alpha(js2_items, check.keys=TRUE)
alpha_js2 <- alpha_js2$total$raw_alpha
alpha_js2 <- round(alpha_js2,2)

freewill_items <- satisfaction_data_usa[, c("fw_1", "fw_2", "fw_3", "fw_4", "fw_5", "fw_6", "fw_7", "fw_8", "fw_9")]
alpha_freewill <- alpha(freewill_items)
alpha_freewill <- alpha_freewill$total$raw_alpha
alpha_freewill <- round(alpha_freewill,2)

jaut1_items <- satisfaction_data_usa[, c("jobaut_3", "jobaut_4", "jobaut_5")]
alpha_jaut1 <- alpha(jaut1_items)
alpha_jaut1 <- alpha_jaut1$total$raw_alpha
alpha_jaut1 <- round(alpha_jaut1,2)

jaut2_items <- satisfaction_data_usa[, c("jobaut2_3", "jobaut2_4", "jobaut2_5")]
alpha_jaut2 <- alpha(jaut2_items)
alpha_jaut2 <- alpha_jaut2$total$raw_alpha
alpha_jaut2 <- round(alpha_jaut2,2)

locus_items <- satisfaction_data_usa[, c("lc1", "lc2", "lc3", "lc4", "lc5", "lc6", "lc7", "lc8", "lc9", "lc10", "lc11", "lc12", "lc13")]
alpha_locus <- alpha(locus_items, check.keys=TRUE)
alpha_locus <- alpha_locus$total$raw_alpha
alpha_locus <- round(alpha_locus,2)

self_eff_items <- satisfaction_data_usa[, c("selfeff_1", "selfeff_2", "selfeff_3")]
alpha_self_eff <- alpha(self_eff_items)
alpha_self_eff <- alpha_self_eff$total$raw_alpha
alpha_self_eff <- round(alpha_self_eff,2)

self_est_items <- satisfaction_data_usa[, c("selfest_1", "selfest_2", "selfest_3", "selfest_4", "selfest_5", "selfest_6", "selfest_7", "selfest_8", "selfest_9", "selfest_10")]
alpha_self_est <- alpha(self_est_items, check.keys=TRUE)
alpha_self_est <- alpha_self_est$total$raw_alpha
alpha_self_est <- round(alpha_self_est,2)

self_con_items <- satisfaction_data_usa[, c("selfcon_1", "selfcon_2", "selfcon_3", "selfcon_4", "selfcon_5", "selfcon_6", "selfcon_7", "selfcon_8", "selfcon_9", "selfcon_10", "selfcon_11", "selfcon_12", "selfcon_13")]
alpha_self_con <- alpha(self_con_items, check.keys=TRUE)
alpha_self_con <- alpha_self_con$total$raw_alpha
alpha_self_con <- round(alpha_self_con,2)

implicit_items <- satisfaction_data_usa[, c("lay_1", "lay_2", "lay_3", "lay_4", "lay_5r", "lay_6r", "lay_7r", "lay_8r")]
alpha_implicit <- alpha(implicit_items, check.keys=TRUE)
alpha_implicit <- alpha_implicit$total$raw_alpha
alpha_implicit <- round(alpha_implicit,2)

reliability_scores <- data.frame(
  Rows = c("Job Satisfaction (T1)", "Job Satisfaction (T2)", "Belief in Free Will (T1)", 
           "Job Autonomy (T1)", "Job Autonomy (T2)", "Locus of Control", 
           "Implicit Beliefs", "Self-Esteem", "Self-Efficacy", "Self-Control"),
  "Reliability Coefficient" = c(alpha_js1, alpha_js2, alpha_freewill, alpha_jaut1, alpha_jaut2, alpha_locus, alpha_self_eff, alpha_self_est, alpha_implicit, alpha_self_con), check.names = FALSE
)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-reliability
#| tbl-cap: Cronbach Alpha Reliability Coefficients for Psychological Constructs measured in Feldman et al. (2018)

kable(reliability_scores)
```
\newpage
## Data Characteristics {#sec-characteristics}

The original research survey data published by @feldman2017 did not require any further data cleaning. Using the aforementioned constructs which are derived from their respective set of survey questions, we were able to replicate all the means and standard deviations, as seen in @tbl-st2summary.
@fig-fwbarst2 shows how in this sample, no participant indicated a free will belief rating less than 3, and that the distribution is strongly left-skewed. The distribution of ages, on the other hand, is right-skewed, with most participants being between 30 and 40 years old (@fig-agebarst2)

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-st2summary
#| tbl-cap: Summary Statistics for Psychological Constructs measured in Feldman et al. (2018)

satisfaction_data_usa_summary <- satisfaction_data_usa %>% 
  summarize(
    mean_age = round(mean(age),2),
    sd_age = round(sd(age),2),
    mean_js1 = round(mean(jobsat1, na.rm = TRUE), 2),
    sd_js1 = round(sd(jobsat1, na.rm = TRUE), 2),
    mean_js2 = round(mean(jobsat2, na.rm = TRUE), 2),
    sd_js2 = round(sd(jobsat2, na.rm = TRUE), 2),
    mean_freewill = round(mean(FWDfwagency, na.rm = TRUE), 2),
    sd_freewill = round(sd(FWDfwagency, na.rm = TRUE), 2),
    mean_jobaut = round(mean(jobaut, na.rm = TRUE), 2),
    sd_jobaut = round(sd(jobaut, na.rm = TRUE), 2),
    mean_jobaut2 = round(mean(jobaut2, na.rm = TRUE), 2),
    sd_jobaut2 = round(sd(jobaut2, na.rm = TRUE), 2),
    mean_locus = round(mean(locus, na.rm = TRUE), 2),
    sd_locus = round(sd(locus, na.rm = TRUE), 2),
    mean_selfcontrol = round(mean(selfcontrol, na.rm = TRUE), 2),
    sd_selfcontrol = round(sd(selfcontrol, na.rm = TRUE), 2),
    mean_selfeff = round(mean(selfeff, na.rm = TRUE), 2),
    sd_selfeff = round(sd(selfeff, na.rm = TRUE), 2),
    mean_selfest = round(mean(selfest, na.rm = TRUE), 2),
    sd_selfest = round(sd(selfest, na.rm = TRUE), 2),
    mean_implicit = round(mean(ess_kind, na.rm = TRUE), 2),
    sd_implicit = round(sd(ess_kind, na.rm = TRUE), 2)
  )

satisfaction_data_usa_summary <- data.frame(
  Variable = c("Age", "Job Satisfaction (T1)", "Job Satisfaction (T2)", "Belief in Free Will (T1)",
               "Job Autonomy (T1)", "Job Autonomy (T2)", "Locus of Control", "Implicit Beliefs",
               "Self-Esteem", "Self-Efficacy", "Self-Control"),
  Mean = c(
          satisfaction_data_usa_summary$mean_age,
           satisfaction_data_usa_summary$mean_js1, 
           satisfaction_data_usa_summary$mean_js2,
           satisfaction_data_usa_summary$mean_freewill,
           satisfaction_data_usa_summary$mean_jobaut,
           satisfaction_data_usa_summary$mean_jobaut2,
           satisfaction_data_usa_summary$mean_locus,
           satisfaction_data_usa_summary$mean_implicit,
           satisfaction_data_usa_summary$mean_selfest,
           satisfaction_data_usa_summary$mean_selfeff,
           satisfaction_data_usa_summary$mean_selfcontrol),
  SD = c(satisfaction_data_usa_summary$sd_age,
        satisfaction_data_usa_summary$sd_js1, 
         satisfaction_data_usa_summary$sd_js2,
         satisfaction_data_usa_summary$sd_freewill, 
         satisfaction_data_usa_summary$sd_jobaut,
         satisfaction_data_usa_summary$sd_jobaut2,
         satisfaction_data_usa_summary$sd_locus,
         satisfaction_data_usa_summary$sd_implicit,
         satisfaction_data_usa_summary$sd_selfest,
         satisfaction_data_usa_summary$sd_selfeff,
         satisfaction_data_usa_summary$sd_selfcontrol))

kable(satisfaction_data_usa_summary)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-fwbarst2
#| fig-cap: Distribution of Free Will Beliefs among Amazon MTurk workers as measured in Feldman et al. (2018)
#| fig-width: 2.5
#| fig-height: 2.5

satisfaction_data_usa_rel %>% ggplot(aes(x=`Belief in Free Will (T1)`)) +
  geom_histogram(binwidth = 0.5, fill = "lightblue", color = "black") +
  xlim(0,6) +
  labs(
       x = "Belief in Free Will Rating",
       y = "Frequency") +
  theme_minimal()
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-agebarst2
#| fig-cap: Distribution of Ages among Amazon MTurk workers as measured in Feldman et al. (2018)
#| fig-width: 2.5
#| fig-height: 2.5

satisfaction_data_usa_rel %>% ggplot(aes(x=`Age`)) +
  geom_histogram(binwidth = 4, fill = "lightblue", color = "black") +

  labs(
       x = "Age",
       y = "Count") +
  theme_minimal()
```
The World Values Survey dataset needed some additional cleaning steps, filtering for certain variables of interest and only for surveys based in the United States of America. See @tbl-wvssummary for summary statistics of selected psychological constructs. Similarly to the MTurk data, free will belief ratings are heavily left-skewed with most respondents indicating high beliefs in free will (@fig-fwbarwvs). The age distribution, however, is much more uniform than in the MTurk data, indicating greater representativeness of the sample (@fig-agebarwvs).

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-wvssummary
#| tbl-cap: "Summary Statistics for selected Psychological Constructs measured in the World Values Survey USA, 1981-2017"

wvs_data_usa_summary <- wvs_data_usa %>% summarise(
  mean_age = round(mean(Age, na.rm = TRUE),2),
    sd_age = round(sd(Age, na.rm = TRUE),2),
  mean_freewill = round(mean(freewill, na.rm = TRUE),2),
  sd_freewill = round(sd(freewill, na.rm = TRUE),2),
  mean_js = round(mean(job_satisfaction, na.rm = TRUE),2),
  sd_js = round(sd(job_satisfaction, na.rm = TRUE),2),
  mean_jaut = round(mean(job_decision, na.rm = TRUE),2),
  sd_jaut = round(sd(job_decision, na.rm = TRUE),2),
  mean_fate = round(mean(fate, na.rm = TRUE),2),
  sd_fate = round(sd(fate, na.rm = TRUE),2),
  mean_pride = round(mean(job_pride, na.rm = TRUE),2),
  sd_pride= round(sd(job_pride, na.rm = TRUE),2),
  )

wvs_data_usa_summary <- data.frame(
  Variable = c("Age","Belief in Free Will", "Job Satisfaction", 
               "Job Autonomy", "Pride in Job", "Belief in Fate"),
  Mean = c(wvs_data_usa_summary$mean_age,
           wvs_data_usa_summary$mean_freewill,
           wvs_data_usa_summary$mean_js,
           wvs_data_usa_summary$mean_jaut,
           wvs_data_usa_summary$mean_pride,
           wvs_data_usa_summary$mean_fate
           ),
  SD = c(wvs_data_usa_summary$sd_age,
         wvs_data_usa_summary$sd_freewill,
         wvs_data_usa_summary$sd_js,
         wvs_data_usa_summary$sd_jaut,
         wvs_data_usa_summary$sd_pride,
         wvs_data_usa_summary$sd_fate
         ))

kable(wvs_data_usa_summary)

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-fwbarwvs
#| fig-cap: "Distribution of Free Will Beliefs among WVS respondents 1981-2022"
#| fig-width: 2.5
#| fig-height: 2.5

wvs_data_usa %>% ggplot(aes(x=freewill)) +
  geom_bar(binwidth = 1, fill = "lightblue", color = "black") +
  labs(
       x = "Belief in Free Will Rating",
       y = "Frequency") +
  theme_minimal()
```
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-agebarwvs
#| fig-cap: Distribution of Ages among among WVS respondents 1981-2022
#| fig-width: 2.5
#| fig-height: 2.5

wvs_data_usa %>% ggplot(aes(x=`Age`)) +
  geom_histogram(binwidth = 4, fill = "lightblue", color = "black") +

  labs(
       x = "Age",
       y = "Count") +
  theme_minimal()
```

\newpage
# Results {#sec-results}

## Amazon Mechanical Turk Sample

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-corst2
#| fig-cap: Correlation Matrix for Psychological Constructs measured in Study 2 of Feldman et al. (2018)
#| fig-width: 10
#| fig-height: 10

satisfaction_data_usa_rel_num <- satisfaction_data_usa_rel %>% dplyr::select(-Age, -Sex, -age_group)
cor_matrix <- cor(satisfaction_data_usa_rel_num, use = 'pairwise.complete.obs')

corrplot(cor_matrix, method = "circle", type='full', tl.col="black", sig.level = 0.05, insig = "blank")

cor_fw_js1 <- cor.test(satisfaction_data_usa_rel$`Belief in Free Will (T1)`, 
                   satisfaction_data_usa_rel$`Job Satisfaction (T1)`, method = "pearson")

cor_fw_js1_correlation_coefficient <- round(cor_fw_js1$estimate,2)
cor_fw_js1_p_value <- cor_fw_js1$p.value
cor_fw_js1_ci <- round(cor_fw_js1$conf.int,2)

cor_fw_js2 <- cor.test(satisfaction_data_usa_rel$`Belief in Free Will (T1)`, 
                   satisfaction_data_usa_rel$`Job Satisfaction (T2)`, method = "pearson")

cor_fw_js2_correlation_coefficient <- round(cor_fw_js2$estimate,2)
cor_fw_js2_p_value <- cor_fw_js2$p.value
cor_fw_js2_ci <- round(cor_fw_js2$conf.int,2)

cor_fw_jaut1 <- cor.test(satisfaction_data_usa_rel$`Belief in Free Will (T1)`, 
                   satisfaction_data_usa_rel$`Job Autonomy (T1)`, method = "pearson")

cor_fw_jaut1_correlation_coefficient <- round(cor_fw_jaut1$estimate,2)
cor_fw_jaut1_p_value <- cor_fw_jaut1$p.value
cor_fw_jaut1_ci <- round(cor_fw_jaut1$conf.int,2)
```

In exploring free will beliefs and their impact on job satisfaction in the context of American Amazon Mechanical Turk workers, our study replicates and extends Feldman et al.'s results. As seen in @fig-corst2, most variables are significantly correlated with each other, with the highest correlations being unsurprisingly between the two measurements of Job satisfaction. Locus of control lacks any significant correlations, which together with its low reliability coefficient designates it as a less decisive control variable. The most important relationship, however, is between free will beliefs and job satisfaction: The belief in free will was significantly positively correlated with job satisfaction both at Time 1 (*r* = `r cor_fw_js1_correlation_coefficient`, *p* \<.001, CI \[`r cor_fw_js1_ci[1]`, `r cor_fw_js1_ci[2]`\]) and at Time 2 (r = `r cor_fw_js2_correlation_coefficient`, *p* \<.001, CI \[`r cor_fw_js2_ci[1]`, `r cor_fw_js1_ci[2]`\]). However, this correlation is not as strong as the one between free will beliefs and job autonomy (*r* = `r cor_fw_jaut1_correlation_coefficient`, *p* \<.001, CI \[`r cor_fw_jaut1_ci[1]`, `r cor_fw_jaut1_ci[2]`\]). This supports the claim that job autonomy may be the most important mediating variable between free will beliefs and job satisfaction. Moreover, a step-wise multiple regression on job satisfaction with all psychological agency constructs (full results in @sec-apx-stepwise) confirmed that the effect of free will beliefs on job satisfaction is robust and the strongest relative impact, even when controlling for other agency constructs (Time 1: *p* \< 0.034; Time 2: *p* \< 0.031).

The replication of study 2 of @feldman2017 supports the claim that there is a direct positive correlation between the belief in free will and job satisfaction over time and controlling for other psychological agency constructs.

### Demographic Factors

@fig-ageboxplotst2 shows that there are slight differences in the median belief in free will among the age groups. However, these differences were not found to be statistically significant, as shown using ANOVA (*F*(4, 204) = 1.886, *p* = .114), even when collapsing individuals over 50 into one age group (*F*(3, 205) = 2.08, *p* = .257). Similarly, using linear regression with Age Group as an interaction term, we do not observe a significant effect of age on the relationship between the belief in free will and job satisfaction.

Looking at sex (@fig-sexboxplotst2), on the other hand, we observe a significant difference in the belief in free will, with women having on average higher ratings of believing in free will (*p* \< 0.05). However, the effect of sex on the relationship between free will beliefs and job satisfaction has not been found to be significant (*p* = 0.58).

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-ageboxplotst2
#| fig-cap: Belief in free will among US-American Amazon Mechanical Turk Workers by Age Group

ggplot(satisfaction_data_usa_rel, aes(x = age_group, y = `Belief in Free Will (T1)`)) +
  geom_boxplot(fill = "lightblue") +
  labs(
       x = "Age Group",
       y = "Belief in Free Will") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  + theme_minimal()
```

```{r, include=FALSE, echo=FALSE}
leveneTest(`Belief in Free Will (T1)` ~ age_group, data = satisfaction_data_usa_rel) # assumption met because p-value is 0.08

anova_result <- aov(`Belief in Free Will (T1)` ~ age_group, data = satisfaction_data_usa_rel)
summary(anova_result)
model <- lm(`Job Satisfaction (T1)`~ `Belief in Free Will (T1)` * age_group, data = satisfaction_data_usa_rel)
summary(model)
t.test(`Belief in Free Will (T1)` ~ Sex, data = satisfaction_data_usa_rel)

lm_interaction <- lm(`Job Satisfaction (T1)` ~ `Belief in Free Will (T1)` * factor(Sex), data = satisfaction_data_usa_rel)
summary(lm_interaction)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-sexboxplotst2
#| fig-cap: Belief in free will among US-American Amazon Mechanical Turk Workers by Sex

ggplot(satisfaction_data_usa_rel, aes(x = factor(Sex, labels = c("Male", "Female")), y = `Belief in Free Will (T1)`)) +
  geom_boxplot(fill = "lightblue") +
  labs(
       x = "Sex",
       y = "Belief in Free Will") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  + theme_minimal()

```

\newpage

## National Sample (WVS)

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-corwvs
#| fig-cap: "Correlation Matrix for selected Psychological Constructs measured in World Values Survey USA 1981-2017"
#| fig-width: 10
#| fig-height: 10

wvs_data_usa_num <- wvs_data_usa %>% dplyr::select(freewill, job_satisfaction, job_decision, job_pride, fate)
wvs_data_usa_num <- wvs_data_usa_num %>% rename("Belief in Free Will" = "freewill",
                                                "Job Satisfaction" = "job_satisfaction",
                                                "Job Autonomy" = "job_decision",
                                                "Pride in Job" = "job_pride",
                                                "Belief in Fate" = "fate",
                                                )
cor_matrix <- cor(wvs_data_usa_num, use = 'pairwise.complete.obs')

corrplot(cor_matrix, method = "circle", type='full', tl.col="black", sig.level = 0.05, insig = "blank")

cor_fw_js1 <- cor.test(wvs_data_usa_num$`Belief in Free Will`, 
                   wvs_data_usa_num$`Job Satisfaction`, method = "pearson")

cor_fw_js1_correlation_coefficient <- round(cor_fw_js1$estimate,2)
cor_fw_js1_p_value <- cor_fw_js1$p.value
cor_fw_js1_ci <- round(cor_fw_js1$conf.int,2)

cor_js_jaut <- cor.test(wvs_data_usa_num$`Job Autonomy`, 
                   wvs_data_usa_num$`Job Satisfaction`, method = "pearson")

cor_js_jaut_correlation_coefficient <- round(cor_js_jaut$estimate,2)
cor_js_jaut_p_value <- cor_js_jaut$p.value
cor_js_jaut_ci <- round(cor_js_jaut$conf.int,2)

cor_fw_jaut <- cor.test(wvs_data_usa_num$`Job Autonomy`, 
                   wvs_data_usa_num$`Belief in Free Will`, method = "pearson")

cor_fw_jaut_correlation_coefficient <- round(cor_fw_jaut$estimate,2)
cor_fw_jaut_p_value <- cor_fw_jaut$p.value
cor_fw_jaut_ci <- round(cor_fw_jaut$conf.int,2)

cor_fw_fate <- cor.test(wvs_data_usa_num$`Belief in Fate`, 
                   wvs_data_usa_num$`Belief in Free Will`, method = "pearson")

cor_fw_fate_correlation_coefficient <- round(cor_fw_fate$estimate,2)
cor_fw_fate_p_value <- cor_fw_fate$p.value
cor_fw_fate_ci <- round(cor_fw_fate$conf.int,2)
```

As shown in @fig-corwvs, the results of the national-level World Values Survey also show a positive correlation between free will beliefs and job satisfaction (*r* = `r cor_fw_js1_correlation_coefficient`, *p* \<.001, CI \[`r cor_fw_js1_ci[1]`, `r cor_fw_js1_ci[2]`\]). Similarly, job autonomy is highly correlated with job satisfaction (*r* = `r cor_js_jaut_correlation_coefficient`, *p* \<.001, CI \[`r cor_js_jaut_ci[1]`, `r cor_js_jaut_ci[2]`\]) and with free will beliefs (*r* = `r cor_fw_jaut_correlation_coefficient`, *p* \<.001, CI \[`r cor_fw_jaut_ci[1]`, `r cor_fw_jaut_ci[2]`\]). Since we created these psychological constructs based on single-item questions of the World Values Survey, internal validity and reliability cannot be guaranteed as concretely as with the Amazon Mechanical Turk survey sample. However, the construct 'Belief in Fate' negatively correlated with the belief in free will, as expected (*r* = `r cor_fw_fate_correlation_coefficient`, *p* \<.001, CI \[`r cor_fw_fate_ci[1]`, `r cor_fw_fate_ci[2]`\]). Similarly, the construct 'Pride in Job' is positively correlated with all other variables except for the belief in fate, as expected.

### Comparison to other countries

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-countries
#| tbl-cap: Correlations Between Belief in Free Will and Job Satisfaction by Country in the World Values Survey

wvs_summary_table <- wvs_filtered %>% group_by(country_code) %>% 
  summarize("FW Mean"=round(mean(freewill),2), n=n())

wvs_filtered_2 <- wvs_filtered %>% filter(js > 0)

correlations <- wvs_filtered_2 %>%
  group_by(country_code) %>%
  do(correlation = cor(.$freewill, .$js, use = "complete.obs"))

correlations <- wvs_filtered_2 %>%
  group_by(country_code) %>%
  summarize(cor_stats = list(cor.test(freewill, js, use = "complete.obs")),
            .groups = 'drop') %>%
  ungroup() %>%
  mutate(correlation = map_dbl(cor_stats, ~ round(.x$estimate, 2)),
         p_value = map_dbl(cor_stats, ~ .x$p.value),
         significance = case_when(
           p_value <= 0.001 ~ "***",
           p_value <= 0.01 ~ "**",
           p_value <= 0.05 ~ "*",
           TRUE ~ ""
         ))

final_table <- left_join(wvs_summary_table, correlations, by = "country_code") %>%
  mutate(`Country Name` = countrycode(country_code, "iso3c", "country.name")) %>%
  dplyr::select(`Country Name`, `FW Mean`, correlation, significance, n) %>%
  rename(`Correlation` = correlation) %>%
  mutate(Correlation = paste0(Correlation, significance)) %>% dplyr::select(-significance) %>% 
  arrange(desc(`FW Mean`))

final_table %>%
  kable("latex", booktabs = T) %>%
  kable_styling(bootstrap_options = "striped") %>%
  footnote(general ="FW Mean = National Average of the Belief in Free Will",
            symbol = "n.s. p > .05. **p < .01. ***p < .001.") %>% 
  row_spec(which(final_table$`Country Name` == "United States"), background = "#ADD8E6")
```

In our analysis, we included data up until 2017 of the World Values Survey (2017), as opposed to only until 2008. This leads to slightly different results than in @feldman2017. Now, the correlations are statistically significant for all countries in the sample and we still observe that the belief in free will and job satisfaction varies similarly to the 2008 data. Compared to the 16 countries sampled in @feldman2017, the United States has a relatively high average belief in free will, coming second just after Mexico (@tbl-countries). While the correlation between the belief in free will and job satisfaction is not as high as in Mexico, it is still one of the stronger significant correlations in the sample of countries (*r* = 0.27, *p* \<.001).

### Demographic Factors

```{r,include=FALSE,message=FALSE,warning=FALSE}

wvs_data_usa$age_group <- cut(wvs_data_usa$Age,
                    breaks = c(18, 30, 40, 50, 60, Inf),
                    labels = c("18-30", "31-40", "41-50", "51-60", "Over 60"),
                    right = FALSE)

# inferential statistics age group and freewill
anova_result <- aov(freewill ~ age_group, data =wvs_data_usa)
summary(anova_result)
model <- lm(job_satisfaction~ freewill * age_group, data = wvs_data_usa)
summary(model)

# inferential statistics sex and freewill
t.test(freewill ~ Sex, data = wvs_data_usa )

lm_interaction <- lm(job_satisfaction ~ freewill * factor(Sex), data = wvs_data_usa)
summary(lm_interaction)
```

Interestingly, there are virtually no differences in the average belief in free will between the different age groups surveyed (@fig-ageboxplotwvs in @sec-apx-box). ANOVA confirmed that there are no significant differences (*F*(4,12809) = 1.032, *p* = 0.358). Age did also not influence the relationship between free will beliefs and job satisfaction (lowest *p* = 0.293). However, there is a significant difference of Sex on free will beliefs (@fig-sexboxplotwvs in @sec-apx-box), with women on average having marginally higher ratings (7.69) than men (7.61) (*p* \< 0.05). We could not observe any significant effect of Sex on the relationship between free will beliefs and job satisfaction (*p* = 0.40).

### Comparison to Amazon Mechanical Turk Sample

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-comparison
#| tbl-cap: Comparative Summary Statistics Between Amazon MTurk and WVS Samples

# Normalize free will scores between the two samples
satisfaction_data_usa_rel_num$freewill_norm <- (satisfaction_data_usa_rel_num$`Belief in Free Will (T1)` - 1)/(6-1)
satisfaction_data_usa_rel_num$jobsatisfaction_norm <- (satisfaction_data_usa_rel_num$`Job Satisfaction (T1)` - 1)/(7-1)

wvs_data_usa$freewill_norm <- (wvs_data_usa$freewill - 1)/(10-1)
wvs_data_usa$jobsatisfaction_norm <- (wvs_data_usa$job_satisfaction - 1)/(10-1)

# Calculate means and sd of those normalized scores

freewill_norm_mean_turk <- round(mean(satisfaction_data_usa_rel_num$freewill_norm, na.rm = TRUE),2)
freewill_norm_sd_turk <- round(sd(satisfaction_data_usa_rel_num$freewill_norm, na.rm = TRUE),2)
jobsat_norm_mean_turk <- round(mean(satisfaction_data_usa_rel_num$jobsatisfaction_norm, na.rm = TRUE),2)
jobsat_norm_sd_turk <- round(sd(satisfaction_data_usa_rel_num$jobsatisfaction_norm, na.rm = TRUE),2)

freewill_norm_mean_wvs <- round(mean(wvs_data_usa$freewill_norm, na.rm = TRUE),2)
freewill_norm_sd_wvs <- round(sd(wvs_data_usa$freewill_norm, na.rm = TRUE),2)
jobsat_norm_mean_wvs <- round(mean(wvs_data_usa$jobsatisfaction_norm, na.rm = TRUE),2)
jobsat_norm_sd_wvs <- round(sd(wvs_data_usa$jobsatisfaction_norm, na.rm = TRUE),2)

comparative_summary_table <- data.frame(
  "Variable" = c("Belief in Free Will", "Job Satisfaction"),
  "Mean" = c(freewill_norm_mean_turk, jobsat_norm_mean_turk),
  "SD" = c(freewill_norm_sd_turk, jobsat_norm_sd_turk),
  "Mean" = c(freewill_norm_mean_wvs, jobsat_norm_mean_wvs),
  "SD" = c(freewill_norm_sd_wvs,jobsat_norm_sd_wvs),
  check.names = FALSE)

comparative_summary_table %>%
  kable("latex", booktabs = T) %>%
  kable_styling(bootstrap_options = "striped") %>% add_header_above(c(" " = 1, "MTurk Sample" = 2, "WVS Sample" = 2))

t_test_result_fw <- t.test(satisfaction_data_usa_rel_num$freewill_norm, wvs_data_usa$freewill_norm)

t_test_result_js <- t.test(satisfaction_data_usa_rel_num$jobsatisfaction_norm, wvs_data_usa$jobsatisfaction_norm)

fwconfint1 <- round(t_test_result_fw$conf.int[1],2)
fwconfint2 <- round(t_test_result_fw$conf.int[2],2)

jsconfint1 <- round(t_test_result_js$conf.int[1],2)
jsconfint2 <- round(t_test_result_js$conf.int[2],2)

```

Since the two samples use different scales to measure belief in free will and job satisfaction, we normalized the values (mapping it onto a scale from 0 to 1), such that we can compare them directly (@tbl-comparison). Results show that the average rating of free will beliefs is higher in the Amazon Mechanical Turk Sample, than in the national sample (*p* \< 0.001, CI \[`r fwconfint1 `, `r fwconfint2 `\]). The opposite trend can be observed in the ratings of job satisfaction, where the national-level WVS sample has on average higher values (*p* \< 0.001, CI \[`r jsconfint1`, `r jsconfint2`\]).

\newpage

# Discussion {#sec-discussion}

## Effect of Free Will Beliefs
The main result of our analysis is that the belief in free will is sufficient to positively influence one's job satisfaction. Both the Amazon Mechanical Turk and WVS data sources showed significant positive correlations between these two psychological constructs. Free will beliefs are found to be  integral to psychological well-being and motivation, and are commonly linked to agency, autonomy, and responsibility [@wang2009]. By fostering an environment that supports these values, organizations can enhance job satisfaction and performance [@stillman2010]. Moreover, we found that job autonomy is the most important factor in mediating the relationship between free will beliefs and job satisfaction. When employees feel they have the freedom to make decisions and control their work processes, this autonomy reinforces their free will beliefs, leading to higher job satisfaction [@goezukara2016]. Our analysis shows that this applies directly to the workplace, since we found a robust effect of free will beliefs on job satisfaction, even when controlling for other psychological agency constructs, suggests a strong and direct relationship. As @clark2017 have found, belief in free will also alleviates distress and anxiety in regards to punishment and moral judgments, which can further positively impact satisfaction at work. This is crucial information, since it can help us effectively design targeted interventions with long-term benefits for employees and organizations [@feldman2017].

## Demographic Variable Effects
Our findings suggest that while there are slight differences in the median belief in free will among different age groups, these differences do not reach statistical significance. This outcome indicates that the belief in free will might be relatively stable across different stages of adult life, not significantly impacted by age-related factors such as life experience and generational values, but rather by cultural factors @feltz2009. Our findings also suggest that a slight differences based on sex, with women reporting higher beliefs in free will. This finding could reflect underlying gender-related socialization processes, where differing societal expectations and roles for men and women might influence their perceptions of autonomy and control [@baumeister1988]. However, even when found significant, this effect was marginal in both samples, suggesting that the influence of free will beliefs on job satisfaction might operate independently of these demographic factors. This independence has strong implications for potential interventions interventions aimed at enhancing employees' sense of autonomy and control could be broadly applicable and beneficial [@sheldon2001].

Nevertheless, unobserved demographic factors could have also contributed to the significant difference in free will beliefs and job satisfaction ratings between the Amazon MTurk sample and the WVS sample. Latter is much more representative due to its larger sample size, more thorough survey methodology, and more equal distribution of ages and occupations. The Amazon MTurk sample was much more restricted, with a much lower median age, and was not cross-occupational in the same sense the more comprehensive WVS sample is. The nature of Amazon MTurk as a novel and mostly low-wage job platform could also explain the comparatively lower job satisfaction rates and higher beliefs in free will.

## Cultural Differences
Our findings indicate that correlations are now statistically significant for all countries in the sample which suggests a pervasive and possibly strengthening relationship between belief in free will and job satisfaction globally. This contrasts with Feldman et al.'s (2018) findings, where the significance of these correlations may have varied across countries. This change could be indicative of a global trend towards greater recognition of the importance of autonomy and personal agency in job satisfaction [@goezukara2016]. Our finding that the United States has a high average belief in free will, second only to Mexico, can be contextualized within the framework of cultural values and individualism. Based on how American cultural norms and values are shaped by freedom, independence, and liberty, the perception and valuation of free will can be expected to be considerably higher than when compared to other countries, influencing the strength of its relationship with job satisfaction. Hofstede's cultural dimensions theory, which explores the impact of a society's culture on the values of its members and how these values relate to behavior, suggests that societies with high individualism scores tend to place a greater emphasis on personal freedom and autonomy, which could contribute to stronger beliefs in free will [@hofstede2001]. Again, this has direct implications for interventions operating within the American context, suggesting a strong focus on such values and constructs to increase overall worker's satisfaction.

## Limitations
Potential weaknesses include the overreliance on self-reporting measures, which are subject to biases; the cross-sectional nature of some studies, which limits causal inference; and the potential for cultural nuances to be oversimplified or to not be fully captured by the statistical measures used.  Additionally, the concept of free will can vary significantly across different cultures. When dealing with constructs that deeply intertwine with personal identity and cultural beliefs it is very important to design survey questions appropriately and make sure to not overstep any boundaries. However, the resulting general nature of these questions could limit capturing the full scope of the concept. The standardized scales used to measure the constructs may have therefore not adequately captured the underlying construct, despite high reliability coefficients.

## Future Directions
The next steps to take would be to further explore societal changes over time and individual differences in the value of free will, which would be possible due to the longitudinal nature of the WVS data. Further research couldalso  investigate the impact of emerging technologies and digital platforms on perceptions of free will, examining how online behavior and the emergence of artificial intelligence influences or reflects these beliefs. Additionally, interdisciplinary approaches combining psychology, sociology, and neuroscience could offer deeper insights into the biological underpinnings and social determinants of free will beliefs. Exploring these areas could significantly enhance our understanding of free will's role in modern society and individual psychology, and help us even more in designing appropriate interventions not only at the workspace but other spheres of our lives.

\newpage

\appendix

# Appendix {#sec-appendix}

## Control Variables Survey Items {#sec-apx-controls}
**Self-Control Scale**

1. I am good at resisting temptation.
2. I have a hard time breaking bad habits. (R)
3. I am lazy. (R)
4. I say inappropriate things. (R)
5. I do certain things that are bad for me, if they are fun. (R)
6. I refuse things that are bad for me.
7. I wish I had more self-discipline. (R)
8. People would say that I have iron self- discipline.
9. Pleasure and fun sometimes keep me from getting work done. (R)
10. I have trouble concentrating.(R)
11. I am able to work effectively toward long-term goals.
12. Sometimes I can’t stop myself from doing something, even if I know it is wrong. (R)
13. I often act without thinking through all the alternatives. (R)

Scale: 1 – Not at all; 5 – Very much.

**Locus of Control**

1.
a. Children get into trouble because their patents punish them too much.
b. The trouble with most children nowadays is that their parents are too easy with
them.
7
2.
a. Many of the unhappy things in people's lives are partly due to bad luck.
b. People's misfortunes result from the mistakes they make.
3.
a. One of the major reasons why we have wars is because people don't take enough
interest in politics.
b. There will always be wars, no matter how hard people try to prevent them.
4.
a. In the long run people get the respect they deserve in this world.
b. Unfortunately, an individual's worth often passes unrecognized no matter how hard
he tries.
5.
a. The idea that teachers are unfair to students is nonsense.
b. Most students don't realize the extent to which their grades are influenced by
accidental happenings.
6.
a. Without the right breaks, one cannot be an effective leader.
b. Capable people who fail to become leaders have not taken advantage of their
opportunities.
7.
a. No matter how hard you try, some people just don't like you.
b. People who can't get others to like them don't understand how to get along with
others.
8.
a. Heredity plays the major role in determining one's personality.
b. It is one's experiences in life which determine what they're like.
9.
a. I have often found that what is going to happen will happen.
b. Trusting fate has never turned out as well for me as making a decision to take a
definite course of action.
10.
a. In the case of the well prepared student there is rarely, if ever, such a thing as an
unfair test.
b. Many times, exam questions tend to be so unrelated to course work that studying is
really useless.
11.
a. Becoming a success is a matter of hard work, luck has little or nothing to do with it.
b. Getting a good job depends mainly on being in the right place at the right time.
12.
a. The average citizen can have an influence in government decisions.
b. This world is run by the few people in power, and there is not much the little guy can
do about it.
13.
a. When I make plans, I am almost certain that I can make them work.
b. It is not always wise to plan too far ahead because many things turn out to be a
matter of good or bad fortune anyhow.

Score one point for response ‘a’ to questions: 1, 3, 4, 5, 10, 11, 12, and 13.

Score one point for response ‘b’ to questions: 2, 6, 7, 8, and 9.

**Self-esteem**

1. I feel that I am a person of worth, at least on an equal basis with others.
2. I feel that I have a number of good qualities.
3. All in all, I am inclined to feel that I am a failure. (R)
4. I am able to do things as well as most other people.
5. I feel that I do not have much to be proud of. (R)
6. I take a positive attitude toward myself.
7. On the whole, I am satisfied with myself.
8. I wish I could have more respect for myself. (R)
9. I certainly feel useless at times. (R)
10. At times I think I am no good at all. (R)
Scale: 1 – Strongly disagree; 7 – Strongly agree.

**Self-efficacy**

1. I am strong enough to overcome life's struggles.
2. At root, I am a weak person. (R)
3. I can handle the situations that life brings.
4. I usually feel that I am an unsuccessful person. (R)
5. I often feel that there is nothing that I can do well. (R)
6. I feel competent to deal effectively with the real world.
7. I often feel like a failure. (R)
8. I usually feel I can handle the typical problems that come up in life.
Scale: 1 – Strongly disagree; 7 – Strongly agree.

**Implicit beliefs**

1. The kind of person someone is, is something very basic about them, and it can't be changed
very much.
2. People can do things differently, but the important parts of who they are can't really be
changed.
3. Everyone is a certain kind of person, and there is not much that they can do to really change
that.
4. As much as I hate to admit it, you can't teach an old dog new tricks. People can't really
change their deepest attributes.
5. Everyone, no matter who they are, can significantly change their basic characteristics. (R)
6. People can substantially change the kind of person they are. (R)
7. No matter what kind of person someone is, they can always change very much. (R)
8. People can change even their most basic qualities. (R)
Scale: 1 – Strongly disagree; 7 – Strongly agree.

## Step-Wise regression results {#sec-apx-stepwise}

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
full_model_t1 <- lm(`Job Satisfaction (T1)` ~ `Belief in Free Will (T1)` + `Self-Esteem` + `Self-Efficacy` + `Self-Control` + `Locus of Control` + `Implicit Beliefs`, data = satisfaction_data_usa_rel)

# Perform stepwise regression using AIC as the criterion
step_model_t1 <- stepAIC(full_model_t1, direction = "both", trace = FALSE)

# Summary of the final model to get coefficients, R-squared, ANOVA, and confidence intervals
summary(step_model_t1)

# ANOVA table for the model
anova(step_model_t1)

# Confidence intervals for the model coefficients
confint(step_model_t1, level = 0.95)

full_model_t2 <- lm(`Job Satisfaction (T2)` ~ `Belief in Free Will (T1)` + `Self-Esteem` + `Self-Efficacy` + `Self-Control` + `Locus of Control` + `Implicit Beliefs`, data = satisfaction_data_usa_rel)

# Perform stepwise regression using AIC as the criterion
step_model_t2 <- stepAIC(full_model_t2, direction = "both", trace = FALSE)

# Summary of the final model to get coefficients, R-squared, ANOVA, and confidence intervals
summary(step_model_t2)

# ANOVA table for the model
anova(step_model_t2)

# Confidence intervals for the model coefficients
confint(step_model_t2, level = 0.95)

```

## Demographic Boxplots for World Values Survey {#sec-apx-box}

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-ageboxplotwvs
#| fig-cap: "Belief in Free Will by Age Group in WVS 1981-2022"

wvs_data_usa %>% drop_na(age_group) %>% ggplot(aes(x = age_group, y = freewill)) +
  geom_boxplot(fill = "lightblue") +
  labs(
       x = "Age Group",
       y = "Belief in Free Will") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  + theme_minimal()
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-sexboxplotwvs
#| fig-cap: "Belief in Free Will by Age Group in WVS 1981-2022"

wvs_data_usa %>% drop_na(Sex) %>% ggplot(aes(x = factor(Sex, labels = c("Male", "Female")), y = freewill)) +
  geom_boxplot(fill = "lightblue") +
  labs(
       x = "Sex",
       y = "Belief in Free Will") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  + theme_minimal()

```

\newpage

# References